\documentclass[12pt]{article}

\usepackage[a4paper, margin=2.5cm]{geometry}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{array}
\usepackage{parskip}
\usepackage{setspace}
\usepackage{microtype}
\usepackage{float}
\usepackage{hyperref}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\setlength{\parindent}{1.5em}
\setlength{\parskip}{0pt}
\onehalfspacing

\hypersetup{
    colorlinks=true,
    urlcolor=black,
    linkcolor=black
}

\begin{document}

%% ---------------------------------------------------------------
%%  TITLE PAGE
%% ---------------------------------------------------------------
\begin{titlepage}
    \centering
    \vspace*{3cm}

    {\LARGE \textbf{Explainable NLP for Caption Similarity and
    Hashtag Recommendation}}\\[1.2em]
    {\large Results (NLP Course)}

    \vspace{3cm}

    \textbf{Group:} A\\[0.8em]

    \textbf{Students:}\\
    Juan Sebastian Pe\~na\\
    Saad Ayomide Olowolayemo\\[0.8em]

    \textbf{Instructor:}\\
    Juan Jos\'e Manjar\'in Colon

    \vfill

    {\large April 2026}
\end{titlepage}

%% ---------------------------------------------------------------
%%  RESULTS
%% ---------------------------------------------------------------
\section*{Results}

\setlength{\parindent}{1.5em}

An initial dataset of 26,623 raw TikTok posts was retrieved from a
PostgreSQL database hosted on Supabase. The data was collected using a
custom TikTok scraper developed in a prior project, publicly available
at \url{https://github.com/PredictiveSocialMedia/Tik-Tok-Recommendation-System}.
During preprocessing, 3,976 posts (14.9\%) were excluded for having fewer
than four non-hashtag characters --- mainly hashtags, URLs, or mentions
without meaningful linguistic content --- leaving a final dataset of 22,647
posts. Engagement metrics showed clear right skew, so all four variables
(likes, comments, views, shares) were log-transformed using
$y' = \log(1 + y)$. Labels were then assigned via a median split of
log-transformed view counts, where the dataset median was 13.06, yielding a
balanced distribution of 11,324 high-engagement and 11,323 low-engagement
posts and a 50\% baseline accuracy for classifier evaluation.

Two complementary representations of caption text were developed. The
first used a TF-IDF matrix ($22{,}647 \times 10{,}000$, density 0.16\%)
to capture lexical patterns via weighted token frequencies, with the
vocabulary restricted to the 10,000 most common terms. The second employed
Sentence-BERT (\texttt{all-MiniLM-L6-v2}) to generate 384-dimensional
semantic embeddings for each caption. UMAP validation reduced the embedding
space to two dimensions, revealing clear clustering and confirming that the
representations capture meaningful semantic structure beyond simple word
overlap. Using both facilitated analysis of lexical and semantic approaches
without presupposing the superiority of either.

Four baseline classifiers were trained on 80\% of the data (18,117 posts)
and evaluated on a stratified held-out set of 4,530 posts: Logistic
Regression with SBERT embeddings, Multinomial Na\"ive Bayes with TF-IDF
features, and KNN with SBERT representations for $k = 5$ and $k = 10$.
All models showed comparable performance, with accuracy ranging from 59\%
to 60\% and macro F1-scores between 0.58 and 0.59
(Table~\ref{tab:classification}).

\begin{table}[H]
    \centering
    \caption{Classification results on the binary engagement label
    (test set, $n = 4{,}530$). Macro F1 averages precision and recall
    equally across both classes.}
    \label{tab:classification}
    \begin{tabular}{llcc}
        \toprule
        \textbf{Model} & \textbf{Feature input} & \textbf{Accuracy}
        & \textbf{Macro F1} \\
        \midrule
        Logistic Regression       & SBERT (384-dim)       & 59\% & 0.59 \\
        Multinomial Na\"ive Bayes & TF-IDF (10{,}000-dim) & 60\% & 0.58 \\
        KNN ($k=5$)               & SBERT (384-dim)       & 59\% & 0.59 \\
        KNN ($k=10$)              & SBERT (384-dim)       & 59\% & 0.58 \\
        \bottomrule
    \end{tabular}
\end{table}

Although overall model performance was similar, Na\"ive Bayes showed class
asymmetry, correctly identifying 79\% of high-engagement posts but only
41\% of low-engagement ones, suggesting a bias toward vocabulary patterns
more common in popular posts. Nevertheless, its overall accuracy was
consistent with the other models.

The comparable performance of SBERT- and TF-IDF-based models suggests the
bottleneck lies in the information content of captions themselves rather
than in the choice of representation. On TikTok, engagement is predominantly
driven by non-textual factors --- audience size, audio, visuals, and platform
algorithms --- which aligns with the 59--60\% accuracy ceiling. Still, the
results suggest captions encode patterns related to the \textit{distribution}
of engagement rather than its exact magnitude, motivating a structural
analysis based on semantic neighborhoods.

For each post, semantic neighborhoods were constructed using cosine
similarity in the SBERT embedding space at three sizes ($K = 5, 10, 20$).
Similarity among nearest neighbors ranged from 0.53 to 0.77, and retrieved
captions conveyed consistent topical content, confirming that the embedding
space captures genuine semantic links rather than surface-level matches.

Within each neighborhood, the variance of the four log-transformed
engagement metrics was calculated and compared to a null distribution from
1,000 random groups of the same size. Variance reduction was defined as
the percentage decrease relative to this random baseline. All twelve values
were positive, ranging from 10.46\% to 13.04\% (Table~\ref{tab:variance}),
showing that semantic neighborhoods consistently exhibited lower within-group
variance than random groups. One-sided Mann-Whitney U tests for each
metric--$K$ combination returned $p \approx 0.0$ across all twelve cases,
strongly indicating that the observed variance reduction was not due to chance.

\begin{table}[H]
    \centering
    \caption{Variance reduction (\%) of semantic neighborhoods relative
    to random groups of equal size. All twelve Mann-Whitney U tests:
    $p \approx 0.0$.}
    \label{tab:variance}
    \begin{tabular}{lcccc}
        \toprule
        $K$ & \textbf{likes\_log} & \textbf{comments\_log}
        & \textbf{views\_log} & \textbf{shares\_log} \\
        \midrule
        5  & 13.04\% & 12.72\% & 12.32\% & 12.46\% \\
        10 & 12.30\% & 12.74\% & 11.36\% & 11.98\% \\
        20 & 11.34\% & 11.27\% & 10.46\% & 10.53\% \\
        \bottomrule
    \end{tabular}
\end{table}

A clear pattern emerged between variance reduction and neighborhood size:
average reduction decreased from approximately 12.6\% at $K=5$, to 12.1\%
at $K=10$, and 10.9\% at $K=20$. Smaller neighborhoods captured only the
most semantically similar posts; larger groups added diversity and raised
variability. This consistent decay across four independent metrics suggests
the effect is geometric in nature --- a consequence of how the SBERT
embedding space is structured --- rather than an artifact of any single
metric. Levene's test provided additional nuance: at $K=5$ the variance
spread between semantic and random groups did not differ significantly
($p > 0.05$ for all metrics), indicating the effect at that scale mainly
reflects a shift in central tendency, whereas at $K=10$ and $K=20$
significant differences in dispersion emerged for most metrics, showing
that structural contrasts between semantic and random groupings grow more
pronounced as neighborhood size increases.

A major finding pertains to hashtag coverage. Initially only 883 posts
(3.9\%) contained hashtags from the bridge-table field. By extracting
hashtags directly from caption text, coverage increased to 100\% ---
an average of 5.58 tags per post across 36,866 unique tags --- enabling
the system to operate across the entire corpus. Recommendations were
generated by aggregating hashtags from the $K$ nearest semantic neighbors
and ranking them by frequency and average log-engagement.

To assess recommendation quality, five representative captions spanning
gaming, DIY crafts, music, comedy, and fitness were passed through the
system; the highest-ranked hashtags consistently matched the topic and tone
of each query. The system was additionally tested on three unseen captions
from the fitness, cryptocurrency, and meal-preparation domains, returning
topically coherent hashtags in each case. Crucially, every recommendation
is traceable to a specific post in the corpus that is semantically similar
to the query, making the system interpretable: suggestions are grounded in
real, identifiable examples rather than generated by an opaque process.

Several limitations must be acknowledged. Caption text captures only part
of what drives engagement, as the 59--60\% accuracy ceiling makes clear.
The SBERT model was primarily trained on English, so performance may degrade
on slang, mixed-language captions, or platform-specific expressions.
Variance reduction also varies with neighborhood size (roughly 10.5\%--13.0\%
depending on $K$), requiring practitioners to calibrate $K$ to their
precision--recall trade-off. Finally, the dataset is a static snapshot;
the KNN index cannot absorb new posts without a full rebuild. These
constraints highlight a key distinction: caption text alone cannot
\textit{predict} engagement precisely, but it does encode meaningful
semantic links to audience response patterns --- and that structural signal
is sufficient for effective, interpretable hashtag recommendations even
without multimodal features.

Future work could address these constraints by adopting multilingual or
domain-adapted language models, exploring incremental indexing to avoid
full rebuilds, and incorporating audio, visual, and metadata signals to
push prediction accuracy beyond the current ceiling. Expanding the dataset
longitudinally and integrating user-level factors would further improve
generalisability.

\end{document}
